---
title: "R Notebook"
output: html_notebook
---
You must write down a short summary of what each plot tells you in order to meet MVP.

Question 1.

Load in the libraries you need, and look at the backpack data. Take note of variable names and dimensions.

```{r}
library(tidyverse)
library(CodeClanData)

backpack <- backpack

dim(backpack)
```

Question 2.

What is the distribution of backpack weights that students are carrying? Choose an appropriate graph to show the distribution.
```{r}
ggplot(backpack)+
  aes(x = backpack_weight)+
  geom_histogram(binwidth = 1, fill = "blue", color = "white")+
  labs(
    title = "\n\nBagpacks weight\n\n",
    x = "\nWeight",
    y = "\nFrequency\n"
  ) +
  scale_y_continuous(breaks = 1:13)+
  scale_x_continuous(breaks = c(2, 4, 6, 8, 10, 12, 14, 16, 18, 20, 22, 24, 26, 28, 30, 32, 34, 36))+
  theme(
    text = element_text(size = 12),
    title = element_text(size = 14),
    axis.text = element_text(size = 10),
    panel.background = element_rect(fill = "white"),
    panel.grid = element_line(colour = "grey90", linetype = "dotted"))

```
# The weight of backpacks carried by students ranged from 2kg to 35kg. The majority of people had bagpacks weighing 8 or 10 kg .

Question 3.

Are there differences in backpack weight between sex?

```{r}
ggplot(backpack)+
  aes(x = sex, y = backpack_weight)+
  geom_dotplot(binaxis = "y", stackdir = "center")+
  scale_y_continuous(breaks = c(2, 4, 6, 8, 10, 12, 14, 16, 18, 20, 22, 24, 26, 28, 30, 32, 34, 36))+
  labs(
    title = "\nBackpack weight by gender\n",
    x = "\nGender\n",
    y = "\n Weight\n"
    )+
  theme(text = element_text(size = 12),
    title = element_text(size = 14),
        axis.text = element_text(size = 10),
    panel.background = element_rect(fill = "white"),
    panel.grid = element_line(colour = "grey90", linetype = "dotted")
    
  )



```


Most females seem to have backpacks between 8-16kg, while most males seem to have backpack weighing 10kg. The heaviest backpack in the data set has been carried by a male.


Question 4.

What is the relationship between body weight and backpack weight?

```{r}

 ggplot(backpack)+
  aes(x = body_weight, y = backpack_weight)+
  geom_point()+
 geom_smooth(method = "lm", se = FALSE, color = "blue")+
  labs(
    title = "\nRelationship bewteen body and backpack weight\n",
    x = "\nBody Weight\n",
    y = "\nBackpack Weight\n"
  )+
   theme(text = element_text(size = 12),
    title = element_text(size = 14),
        axis.text = element_text(size = 10),
    panel.background = element_rect(fill = "white"),
    panel.grid = element_line(colour = "grey90", linetype = "dotted")
    
  )
```

There doesn't seem to be a clear relationship between body and backpack weight.
Question 5.

Is this relationship different between men and women? Choose an appropriate plot and interpret your findings.
```{r}
ggplot(backpack)+
  aes(x = body_weight, y = backpack_weight, col = sex)+
  geom_point()+
   geom_smooth(method = "lm", se = FALSE)
+
facet_wrap(~sex)
```
Females tended to have lower weight (below 150 lb) but were carrying backpacks weighing between 8-16kg/lb, while males weighed mainly over 150lb and were carrying backpacks with various weight.


What proportion of students report back pain?
```{r}
ggplot(backpack)+
  aes(x = factor(back_problems))+
     geom_bar()
```
Around one third of students seemed to report back problems.
```{r}
ggplot(backpack)+
  aes(x = body_weight, fill = sex)+
  geom_bar()+
   geom_smooth(method = "lm", se = FALSE)
+
facet_wrap(~sex)
```

Question 7.

Is this proportion different between men and women?
```{r}
ggplot(backpack)+
  aes(x = factor(back_problems), fill = sex)+
     geom_bar()
```

Mostly females reported back problems.

Question 8.

Now you’ve investigated your data separately, create plots to describe the relationship between backpack weight, body weight, gender and reported backpain.
```{r}
ggplot(backpack)+
  aes( x = body_weight, y = backpack_weight, group = back_problems, col = sex )+
  geom_point()+
  geom_smooth(method = "lm",se = FALSE, color = "red")+
  facet_grid(~back_problems~sex)
```

