---
title: "R Notebook"
output: html_notebook
---


1. Load the housing_prices.csv data set and undertake an initial exploration of the data. You will find details on the data set on the relevant Kaggle page.

```{r}
library(tidyverse)
library(GGally)
library(skimr)
library(fastDummies) # to make dummy variables
library(modelr)
library(ggfortify)
library(fastDummies)
library(janitor)
housing_prices <- read_csv("data/housing_prices.csv")

housing_prices %>% select(total_rooms, total_bedrooms)
skim(housing_prices) 

# there were 200 misssing values in total bedrooms  and I decided to leave them for now


summary(housing_prices)

housing_dummy <- housing_prices %>%  dummy_cols(select_columns = "ocean_proximity", remove_first_dummy = FALSE, remove_selected_columns = TRUE) %>% select(-ocean_proximity_ISLAND)
housing_dummy <- clean_names(housing_dummy)


```

2. We expect the total_rooms of houses to be strongly correlated with total_bedrooms. Use ggpairs() to investigate correlations between these two variables.

```{r}
housing_dummy %>% ggpairs(columns = c("total_rooms","total_bedrooms"))


```
3. So, we do find significant correlations. Let’s drop total_bedrooms from the dataset, and use only total_rooms going forward.
```{r}
housing_dummy <- housing_dummy %>% select(-total_bedrooms)
housing_dummy
```
4. We are interested in developing a regression model for the median_house_value of a house in terms of the possible predictor variables in the dataset.

4a. Use ggpairs() to investigate correlations between median_house_value and the predictors (this may take a while to run, don’t worry, make coffee or something).

```{r}
housing_dummy %>% ggpairs(columns = c("median_house_value", "housing_median_age", "total_rooms", "population", "households", "median_income", "ocean_proximity_1h_ocean", "ocean_proximity_inland", "ocean_proximity_near_bay", "ocean_proximity_near_ocean"))
# mod1 <- lm(median_house_value ~ total_rooms + population + households + median_income + housing_median_age, housing_prices)
# autoplot(mod1)
```
Strong correlation between median_house_value and median income. There is also a moderate correlation between median_house_value and ocean_proximity_inland.
There is also a very strong correlation between household and population, total rooms and households, and also between population and total rooms, but we are not interested in these ones for now.

4b. Perform further ggplot visualisations of any significant correlations you find.

```{r}
housing_dummy %>% 
  ggplot(aes(median_house_value, median_income))+
  geom_point()+
  geom_smooth(method = "lm")

# housing_dummy %>% 
#   ggplot(aes(median_house_value, ocean_proximity_inland))+
#   geom_col()
```
Shortly we may try a regression model to fit the categorical predictor ocean_proximity. Investigate the level of ocean_proximity predictors. How many dummy variables do you expect to get from it?
```{r}
# housing_prices %>% group_by(ocean_proximity) %>% 
#   summarise(proximity_type_count = count(ocean_proximity))
```
 Ocean_proximity has 5 levels which indicates that there could be 4 dummy variables (the fifth one would be dropped)

6. Start with simple linear regression. Regress median_house_value on median_income and check the regression diagnostics.

```{r}

model <- lm(formula = median_house_value ~  median_income, data =  housing_dummy)
model
autoplot(model)

```

For each 10,000 USD the median house value goes by 39987 USD.

```{r}
mod2 <- lm(median_house_value ~ median_income + total_rooms, housing_prices)
mod2


autoplot(mod2)
```

