---
title: "Clustering homework"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

<br>

This homework involves the `mall_customers.csv` data we've provided, based on data collected in a shopping mall. It will give you an opportunity to practice your new clustering skills.


```{r}

library(tidyverse)
# read in data

customers <- janitor::clean_names(read_csv("data/mall_customers.csv"))

customers

```

<br>

# MVP

Task 1

* Do some exploratory analysis on the data. Comment on findings. 

```{r}
skimr::skim(customers)
head(customers)


customers %>% 
  ggplot(aes(x= spending_score_1_100, y = annual_income_k, colour = gender)) + 
  geom_point()

```
There are five variables in the data set. There seem to be more females than males in the data set. There is more people aged 40-50 in the dataset, with teh peak of earnings around 50k - 80k.

```{r message = FALSE}
GGally::ggpairs(customers)
```


<br>

Task 2

* We are interested in creating a marketing campaign to target customers based on their spending score and annual income. Select an appropriate value for k and perform k-means clustering to find if there are meaningful clusters in the data to target the customers. 

```{r}

customers_new <- customers %>% 
  select(c(annual_income_k, spending_score_1_100))

head(customers_new)

```
```{r}
# we need to scale it

customers_new_scale<- customers_new %>% 
  mutate_if(is.numeric, scale)

customers_new_scale

# check if it worked 

customers_new_scale %>% 
  as_tibble() %>% 
  pivot_longer(cols = c(annual_income_k, spending_score_1_100),
               names_to = "type", 
               values_to = "value",
               ) %>% 
  group_by(type) %>% 
  summarise(mean = round(mean(value)),
            sd = sd(value))
```

```{r}
# let's check how many clusters are optimal but first we need to set min and max of clusters to look at 

library(broom) # needed for tidy()

max_k <- 10

k_clusters <- tibble(k = 1:max_k) %>% 
 mutate(kclust = map(k, ~kmeans(customers_new_scale, .x, nstart = 25)),
        tidied = map(kclust, tidy),
        glanced = map(kclust, glance),
        augmented = map(kclust, augment, customers_new)
        ) 
 
 
# an elbow method:
# clusterings <- k_clusters %>% 
#   unnest(glanced)
# 
# clusterings
# 
# # we can visualise the results
# 
# clusterings %>% 
#   ggplot(aes(x = k, y = tot.withinss))+
#   geom_point()+
#   geom_line()+ scale_x_continuous(breaks = seq(1, 20, by = 1))

# potentially k = 3
library(factoextra)

fviz_nbclust(customers_new_scale,
             kmeans, 
             method = "wss",
             nstart = 30)           

# silhouette method


fviz_nbclust(customers_new_scale,
             kmeans, 
             method = "silhouette",
             nstart = 30)
 # gapstatistic method


fviz_nbclust(customers_new_scale,
             kmeans, 
             method = "gap_stat",
             nstart = 30,
             k.max = 10)



```


I am going to go with k = 5

```{r}
# let's perform K-means clustering
clustered_customers <- kmeans(customers_new_scale,
                              centers = 5, nstart = 30
                              )

clustered_customers
```

```{r}
tidy(clustered_customers, 
     col.names = colnames(customers_new_scale))

customers_new_aug <- augment(clustered_customers, customers_new)
```


<br>

Task 3

* Visualise the clustering for your chosen value of k.

```{r}
customers_new_aug %>% 
  ggplot(aes(x = annual_income_k, y = spending_score_1_100, colour = .cluster))+
  geom_point()
```


<br>

Task 4

* Do you think the clustering seems a good fit for this data?

Yes, it makes sense.

<br>

Task 5

* Comment on the attributes of one or two of the clusters (maybe even give them an appropriate label if you like - similar to section 4.1 of the 'Segmentation & clustering intro' lesson where we saw labels such as "Super Fans" and "Tag-Alongs"). 

```{r}
tidy(clustered_customers)
```
Cluster 1 are the people who earn more and spend less,
Cluster 2 are those, who earn high but also spend high,
Cluster 3 are those who are average earners and average spenders,
cluster 4 are those who are on a lower end of an earning scale but their spending is high, 
Cluster 5 are those who earn little but also spend little.






