---
title: "R Notebook"
output: html_notebook
---
Question 1

Tidy up the data ready for regression:

```{r}
library(tidyverse)
library(modelr)

# visualisation packages
library(GGally)
library(ggfortify)

kc_house <- read_csv("data/kc_house_data.csv")
# remove columns suggested as not useful
kc_house_tidy <- kc_house %>% 
  select(-c(id, date, sqft_living15, sqft_lot15, zipcode, long, lat)) %>% 
  mutate(renovated = as.logical(yr_renovated),
         renovated = as.numeric(renovated),
         view = as_factor(view),
         condition = as_factor(condition),
         grade = as_factor(grade))
         
kc_house_tidy %>% 
  distinct(grade)
```

Question 2
Check for aliased variables using the alias() function (this takes in a formula object and a data set). [Hint - formula price ~ . says ‘price varying with all predictors’, this is a suitable input to alias()]. Remove variables that lead to an alias. Check the ‘Elements of multiple regression’ lesson for a dropdown containing further information on finding aliased variables in a dataset.

```{r}
alias(price ~ ., kc_house_tidy)

# sqft_living = sqft_above + sqft_basement) - these are aliased variables, so will remove sqft_basement
kc_house_tidy <- kc_house_tidy %>% 
  select(-sqft_basement)

alias(price ~ ., kc_house_tidy)
```
Question 3
Systematically build a regression model containing up to four main effects (remember, a main effect is just a single predictor with coefficient), testing the regression diagnostics as you go * splitting datasets into numeric and non-numeric columns might help ggpairs() run in manageable time, although you will need to add either a price or resid column to the non-numeric dataframe in order to see its correlations with the non-numeric predictors.

```{r}
houses_tidy_numeric <- kc_house_tidy %>% 
  select(-yr_renovated) %>% 
  select_if(is.numeric)

houses_tidy_nonnumeric <- kc_house_tidy %>%
  select_if(function(x) !is.numeric(x))

houses_tidy_nonnumeric$price <- kc_house_tidy$price

ggpairs(houses_tidy_numeric, progress = FALSE)
# saves teh plot for future reference, can also provide dimensions to make it bigger
ggsave("ggpairs_numeric.png", width = 15, height = 15)
ggpairs(houses_tidy_nonnumeric, progress = FALSE)

ggsave("ggpairs_nonnumeric.png", width = 15, height = 15)
```

 Based on the above I will build models with the following variables: sqft_living, bathrooms, view and grade.
 
```{r}
mod1 <- lm(price ~ sqft_living, kc_house_tidy)
summary(mod1)
autoplot(mod1)

# example fo checking with anova

mod2 <- lm(price~sqft_living + grade, kc_house_tidy)
anova(mod1, mod2)

# example with grouping

houses_tidy <- kc_house_tidy %>% 
  mutate(grade_grouped = as.factor(if_else(as.numeric(grade) < 8, 7,as.numeric(grade) )))
         
         
```

```{r}
plot(mod2)
```

```{r}
mod2 <- lm(price ~ bathrooms, kc_house_tidy)
summary(mod2)
autoplot(mod2)
```
```{r}
# check the price ~ view
mod3 <- lm(price ~ view, kc_house_tidy)
summary(mod3)
autoplot(mod3)
```
```{r}
# check price~grade
mod4 <- lm(price~ grade, kc_house_tidy)
summary(mod4)
autoplot(mod4)
```
```{r}
# create a new data set called houses_remaining_resid
houses_remaining_resid_num <- houses_tidy_numeric %>% 
  add_residuals(mod1) %>% 
  select(-c("sqft_living", "price"))

houses_remaining_resid_num %>% 
  ggpairs()
```

```{r}
mod6a <- lm(price ~ sqft_living + view, kc_house_tidy)

summary(mod6a)
autoplot(mod6a)
anova(mod1, mod6a)

```
 
 
```{r}

mod6b <- lm(price ~ sqft_living + bathrooms, kc_house_tidy)
summary(mod6b)
autoplot((mod6b))
anova(mod6b, mod1) # mod6b not worth keeping
```
 
```{r}
houses_remaining_resid <- kc_house_tidy %>% 
  add_residuals(mod6a) %>% 
  select(-c("sqft_living", "price", "grade"))
houses_remaining_resid %>% 
  ggpairs(transform = FALSE)


houses_model <- kc_house_tidy
```
 
So we have a model with the first predictor being sqft_living and second view 

```{r}
mod7a <- lm(price ~ sqft_living + view + grade, kc_house_tidy)
summary(mod7a)
autoplot(mod7a)
anova(mod6a, mod7a)
```

```{r}
mod7b <- mod7a <- lm(price ~ sqft_living + view + bathrooms, kc_house_tidy)
summary(mod7b)
autoplot(mod7b)
anova(mod6a, mod7b)
```

grade (especially 10-12) seem to have a greater significance than the number of bathrooms



