---
title: "R Notebook"
output: html_notebook
---
```{r}
library(tidyverse)
library(janitor)
library(gtools)
titanic <- read_csv("data/titanic.csv")
source("../../../../codeclan_work/week _06/day_1/prob.R")
```

```{r}
titanic_new <- titanic %>% mutate(crew = if_else(class %in% c("1st", "2nd", "3rd"), FALSE, TRUE))

```
##  1.1 The Titanic

Calculate the conditional probabilities prob(survived∣crew) and prob(survived∣passenger)to settle this dispute.

Have a look at the distinct values of the class variable, this will tell you whether each individual was a passenger (1st, 2nd or 3rd class) or a member of the crew (treat restaurant staff as crew)

Was meant to calculate against the crew not everybody on the boat
Answer:
Probability of crew surviving is 0.09.

```{r}
titanic_new %>% filter(crew == TRUE & survived == "yes") %>% 
  summarise(p = n()/nrow(titanic_new))
```
Answer: Probability of passenger surviving is 0.226.

```{r}

titanic_new %>% filter(crew == FALSE & survived == "yes") %>% 
  summarise(p = n()/ nrow(titanic_new))
```
1.2 Probability Consolidation Questions

Question 1.
If I roll a six-sided dice once, what is the probability of getting a number less than 5?

event A: roll one of {1, 2, 3, 4}
Answer: Prob(A) 4/6 = 2/3 = 0.6667

```{r}
rolldie(1)
```

Question 2.
How many outcomes are there in the sample space of an experiment in which we toss three dice?

```{r}
s_three_dice <- rolldie(3)
s_three_dice
```
Answer: 216 outcomes

Question 3

In an experiment where we roll four dice:

- How many outcomes are in the event where the total of the four dice equals 5
, 6 or 7)?

1,1,1,2 = 5

1, 1, 2, 2 = 6
1, 1, 1, 3 = 6

1, 1, 1, 4 = 7
1, 1, 2, 3 = 7
1, 2, 2, 2 = 7

Answer: 6 different outcomes

```{r}
s_four_dice <- rolldie(4)
s_four_dice

s_four_dice %>% filter(X1+X2+X3+X4 == 5 | X1+X2+X3+X4 == 6 | X1+X2+X3+X4 == 7) %>% 
  summarise(p = n()/nrow(s_four_dice))

```

what is the probability of this event?

Answer: Probability of throwing 4 dice, where the sum equals 5, 6 or 7 is 0.0262.


what assumptions do we apply in answering part ii?

we're assuming that order in which numbers are thrown doesn't matter.
Hint
For part i, rowSums(.) may help you setup up the necessary random variable total (or you can use pivot_longer())

```{r}
s_four_dice %>% filter(rowSums(.) %in% c(5, 6, 7)) %>% 
   summarise(p = n()/nrow(s_four_dice))
```
Question 4.
Five chairs are placed in a row in a room, along with five people of differing height. The people sit down randomly on the chairs. What is the probability that they sit in order of decreasing height, left to right?

Hint
This is an arrangement where order matters. Of all the possible arrangements, how many are correct?

```{r}
height <- c( 1.2, 1.3, 1.4, 1.5, 1.6)

sitting_order <- permutations(n = 5, r = 5, v = height, repeats.allowed = FALSE) %>% 
  as_tibble()

sitting_order %>% unite(height, c(V1:V5), sep = "_", remove = TRUE) %>% 
filter(str_detect(height,"1.6_1.5_1.4_1.3_1.2")) %>% 
  summarise(n()/nrow(sitting_order))
  
 

```
Answer: The probability is 1/120 = 0.0083

Question 5:
```{r}
s_six_dice <- rolldie(6)
```


I roll six dice:

What is the probability I see the numbers 6, 5, 4, 3, 2 and 1 in order on the dice, left to right?

1/46656




```{r}

s_six_dice %>% filter(isin(., c(6:1), ordered = TRUE)) %>% 
  summarise(p = n()/nrow(s_six_dice))
```
what is the probability I see one each of the numbers 6, 5, 4, 3, 2 and 1 on the dice regardless of order?

```{r}
s_six_dice %>% filter(isin(., c(6:1), ordered = FALSE)) %>% 
  summarise(p = n()/nrow(s_six_dice))
```
answer: 0.0154321

Question 6.
If I draw a single card from a 52-card deck, and it is black, what is the probability that its suit is spades?

```{r}
# 52 card deck - half of it is black, quarter of it are spades

cards()

```




Hints
Investigate the cards() function in the prob.R file, if you’d like to try this in R (make a deck without jokers).
You could think about adding your own black_card random variable to the sample space, which would be TRUE if the suit is ‘Spade’ or ‘Club’, and FALSE otherwise.

