---
title: "R Notebook"
output:
  html_document:
    df_print: paged
---

```{r}
library(tidyverse)

```

```{r}
movies <- read_csv("data/movies.csv")
```
```{r}
view(movies)
```

Q1
- order them by release date, budget, popularity etc.
- see the average budget for a specific year or length of the movie
- see the average length of the movie for a specific year
- how many or what movies do not have a website, is it connected with the length of the movie, populatity, budget, revenue etc
- are there any movies with the original_title being different to the title

- see the relationship between vote average and vote count

Q2
```{r}
movies_selected <- movies %>%
  select(title, runtime, budget)

movies_selected
  
```
Q3 Ok, now take your subsetted data movies_selected, and count the number of missing values in each column.
```{r}
movies_selected %>% 
mutate(missing = (is.na(movies_selected))) %>% 
summarise(total_missing = sum(missing))
```
Q4
here’s not too many NAs, but there are actually quite a few movies with a runtime of 0 in this dataset. Since it’s impossible for a movie to have a runtime of 0, use the function na_if to convert runtimes of length 0 into NAs. Save this into a new object called movies_runtime. Then count how many missing values there are now in the column runtime.

```{r}
movies_runtime <- movies_selected %>% 
  select(runtime) %>% 
  mutate(runtime = na_if(runtime, 0)
  )

movies_runtime %>% 
  filter(is.na(movies_runtime))
```
Q5
Impute the missing values in runtime with the median runtime using coalesce and save this into a new object called movies_imputed. There should now be no missing values in the column runtime - check this to make sure.


```{r}
movies_imputed <- movies_runtime %>% 
  mutate(runtime = coalesce(runtime, median(runtime), na.rm = TRUE))

movies_imputed %>% 
  filter(is.na(movies_imputed))
```

Q6
Finally, use the slice_min and slice_max functions on movies_imputed to get the movies with the 10 shortest and 10 longest runtimes.


```{r}
movies_imputed %>% 
  slice_min(movies_imputed)


```
```{r}
movies_imputed %>% 
  slice_max(movies_imputed
            )
```



Q7
Using the if_else function, impute anything with a budget below $100 with the median budget. Overwrite your movies_imputed so it contains the new budget values.

```{r}
movies_imputed <-movies_selected %>% 
  select(budget) %>% 
  mutate(budget = if_else(budget < 100, median(budget), budget))
  
movies_imputed %>% 
 arrange(movies_imputed)
  
```

