---
title: "R Notebook"
output: html_notebook
---

Task 1.
Load the data again, clean_names(), and re-familiarise yourself with it


```{r}
library(tidyverse)
library(janitor)
library(skimr)
library(infer)
ames <- clean_names(read_csv("data/ames.csv"))
skim(ames)

```

Task 2.
Investigate the distribution of lot_area. Is the distribution roughly normal? If not, what problems do you find?
```{r}
ames %>% 
  ggplot(aes(x = lot_area))+
  geom_histogram(colour  = "white", bins = 100)

ames %>% 
  summarise(
   mean = mean(lot_area),
   median = median(lot_area)
  )
```
This distribution looks slightly skewed towards right. The difference between mean and median looks moderate.

Task 3.
Compute and visualise a bootstrap sampling distribution for the mean(lot_area) of the sold houses.
one way:

```{r}
# bootstrap the sample


bootstrap_resample_ames <- ames %>%
  rep_sample_n(size = 2930, replace = TRUE, reps = 10000) %>% 
  summarise(mean_lot_area = mean(lot_area))
  
bootstrap_resample_ames

# let's plot it

bootstrap_resample_ames %>% 
  ggplot(aes(x=mean_lot_area)) +
  geom_histogram(fill = "lightgreen", colour = "white")
```
  

another way:

```{r}
infer_resample_ames <- ames %>% 
  specify(response = lot_area) %>% # specify the variable of interest
  generate(reps = 10000, type = "bootstrap") %>% # generate the replicate resamples
  calculate(stat = "mean")
infer_resample_ames %>% 
  ggplot(aes(x = stat))+
  geom_histogram(fill = "lightgreen", colour = "white")
```

Task 4.
Use your bootstrap distribution to calculate a 95%
 CI for mean(lot_area), and visualise it on the distribution 

```{r}

```
```{r}

# calculate Confidence intervals 
lot_area_ci_95 <- infer_resample_ames %>% 
  get_confidence_interval(level = 0.95, type = "percentile")

lot_area_ci_95
```
visualise:

```{r}
infer_resample_ames %>% 
  visualise(bins = 20) +
  shade_confidence_interval(endpoints = lot_area_ci_95)

```

Task 5.
You would like to know the mean(lot_area) of the sold houses with higher confidence. Calculate the 99%
 CI for this variable (you can re-use your bootstrap distribution from above). Is it narrower or broader than the 95%
 CI? Does that make sense?
 
```{r}
lot_area_ci_99 <- infer_resample_ames %>% 
  get_confidence_interval(level = 0.99, type = "percentile")

# visualise

infer_resample_ames %>% 
  visualise(bins = 20) +
  shade_confidence_interval(endpoints = lot_area_ci_99)
```
 
 Broader, Yes it makes sense as if CI is higher then it means that 99% of all values has to fall within the top and bottom bound.

Task 6.
Calculate the point estimate of the mean(lot_area)
```{r}

# point estimate
mean_lot_area <- infer_resample_ames %>% 
  summarise(
    mean = mean(stat))
mean_lot_area
```
 Extension


Task 1.
Calculate a point estimate and 95%
 CI for the proportion of houses in the data built before 1920. Does the number of reps you use matter? [Investigate reps from 200
 up to 50000
, memory of your laptop permitting].
## 200 reps
```{r}
infer_resample_before1920 <-  ames %>% 
filter(year_built < 1920) %>% 
  specify(response = lot_area) %>% # specify the variable of interest
  generate(reps = 200, type = "bootstrap") %>% # generate the replicate resamples
  calculate(stat = "mean")

infer_resample_before1920 %>% 
  visualise(bins = 20) +
  shade_confidence_interval(endpoints = lot_area_ci_95)
  
```
##500 reps

```{r}
infer_resample_before1920 <- ames %>% 
filter(year_built < 1920) %>% 
  specify(response = lot_area) %>% # specify the variable of interest
  generate(reps = 500, type = "bootstrap") %>% # generate the replicate resamples
  calculate(stat = "mean")

infer_resample_before1920 %>% 
  visualise(bins = 25) +
  shade_confidence_interval(endpoints = lot_area_ci_95)
```
##1000 reps
```{r}
infer_resample_before1920 <- ames %>% 
filter(year_built < 1920) %>% 
  specify(response = lot_area) %>% # specify the variable of interest
  generate(reps = 1000, type = "bootstrap") %>% # generate the replicate resamples
  calculate(stat = "mean")

infer_resample_before1920 %>% 
  visualise(bins = 25) +
  shade_confidence_interval(endpoints = lot_area_ci_95)
```
## 10000
```{r}
infer_resample_before1920 <- ames %>% 
filter(year_built < 1920) %>% 
  specify(response = lot_area) %>% # specify the variable of interest
  generate(reps = 10000, type = "bootstrap") %>% # generate the replicate resamples
  calculate(stat = "mean")

infer_resample_before1920 %>% 
  visualise(bins = 25) +
  shade_confidence_interval(endpoints = lot_area_ci_95)
```
##25000 reps

```{r}
infer_resample_before1920 <- ames %>% 
filter(year_built < 1920) %>% 
  specify(response = lot_area) %>% # specify the variable of interest
  generate(reps = 25000, type = "bootstrap") %>% # generate the replicate resamples
  calculate(stat = "mean")

infer_resample_before1920 %>% 
  visualise(bins = 25) +
  shade_confidence_interval(endpoints = lot_area_ci_95)
```
##35000 reps

```{r}
infer_resample_before1920 <- ames %>% 
filter(year_built < 1920) %>% 
  specify(response = lot_area) %>% # specify the variable of interest
  generate(reps = 35000, type = "bootstrap") %>% # generate the replicate resamples
  calculate(stat = "mean")

infer_resample_before1920 %>% 
  visualise(bins = 25) +
  shade_confidence_interval(endpoints = lot_area_ci_95)
```
##45000 reps
```{r}
infer_resample_before1920 <- ames %>% 
filter(year_built < 1920) %>% 
  specify(response = lot_area) %>% # specify the variable of interest
  generate(reps = 45000, type = "bootstrap") %>% # generate the replicate resamples
  calculate(stat = "mean")

infer_resample_before1920 %>% 
  visualise(bins = 25) +
  shade_confidence_interval(endpoints = lot_area_ci_95)
```
## 50000 reps
```{r}
infer_resample_before1920 <- ames %>% 
filter(year_built < 1920) %>% 
  specify(response = lot_area) %>% # specify the variable of interest
  generate(reps = 50000, type = "bootstrap") %>% # generate the replicate resamples
  calculate(stat = "mean")

infer_resample_before1920 %>% 
  visualise(bins = 25) +
  shade_confidence_interval(endpoints = lot_area_ci_95)
```

